# Project initiation {#intro}

```{r}
```

[ To do in this section: define a minimum package of resources to get basic data for a given basin]

## Concepts

Add text about site choice, capacity and resource assessment.

After sel-assessment, one of the following scenario should be identified as the most suitable to describe the project:

**Scenario 01: low resources** – In this alternative, low financial resources are assumed. Methods are chosen based on cost. A medium to high level of capacity is assumed.

**Scenario 02: low uncertainty** – This scenario aims at reducing the uncertainty of sampling and analysis. Nevertheless a trade-off with cost and resources needs to be found.

**Scenario 03: Participation of local communities** Simpler methods involving citizen science and monitoring are used. More frequent, cheaper, less accurate monitoring and sampling are preferred.


## Tools


**River basin information**\

- [Hydrosheds](https://hydrosheds.org)


**Capacity assessment**\

- [A Brief Review of 20 Tools to
Assess Capacity](https://www.humanitarianlibrary.org/sites/default/files/2014/02/capacity_development_group_undp_2005_brief_review_20_tools_to_assess.pdf)

- [Integrated Water Resources Management, as measured by SDG indicator 6.5.1](http://iwrmdataportal.unepdhi.org)


**Assessing availabe resources**\

- Research on data costing could provide a tool to assess available budget


## An example
The study was conducted on the Akaki River basin (Awash Wenz 2 in Hydrosheds REF). The catchment is located in central Ethiopia along the western mar- gin of the Main Ethiopian Rift and geographically located between 8°46´–9°14´N and 38°34´–39°04´E, covering an area of about 1500 km2. 

It experiences a temperate Afro-Alpine climate with daily average temperatures from 9.9 to 24.6 °C and annual mean rainfall is 1254 mm, as measured at Addis Ababa Observatory .Three main seasons can be observed: the major rainy one, locally known as Kiremt from June to September, contributing about 70% of the total annual rainfall. A minor rainy season, locally known as Belg, from mid-February to mid-April. The remaining five months are dry season (Ethiopian Public health Institute, 2017).

To make the case study more manageable, only one Water Security dimension will be considered: water quality.

```{r, message= FALSE, echo = FALSE, warning = FALSE, out.width = "100%"}
library(leaflet)

m <- leaflet() %>%
  addTiles() %>%addProviderTiles("Stamen.Terrain")%>%
  setView(lng = 38.7, lat = 9, zoom = 12)# Add default OpenStreetMap map tiles
m

```
